\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-09-18 notes/2loopNLSM.tex
Added a few references

Color-kinematics duality is a statement about the off-shell graphical simplicity that manifest redundancy in the on-shell observables of gauge and gravity theories. Ignoring factors of $i$ and the coupling constant, a generic $m$-point $L$-loop amplitude in a color-dual theory -- such as NLSM, YM, or bi-adjoint scalar theory -- can be written as
\be
\mathcal{A}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int \frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{c_\Gamma n_\Gamma}{d_\Gamma} .
\ee
Here the spacetime dimension is $D$.
Even though the underlying theory may have quartic (or higher multiplicity) vertices, it is possible to multiply and divide by propagators so that the sum runs over only purely cubic graphs $\Gamma$.
Overcounting is removed by the internal symmetry factor $S_\Gamma$, which tracks the number of graph symmetries with external legs held fixed.
The color factor $c_\Gamma$ is the string of structure constants $f^{abc}$ associated with the graph $\Gamma$,\footnote{This paper is focused on theories with particles that transform in the adjoint.  However, the double copy can be applied to a far more diverse set of theories with particles that transform in the fundamental or more diverse representations, see for example \cite{Johansson:2017srf, Johansson2014zca, Johansson:2015oia, Johansson:2019dnu} ref[Ingrid+JJ, Aslan+JJ] \jm{I cited DF$^2$ and fundamental fermions. Is there anything else anyone wants to mention?}.}
the denominator $d_\Gamma$ is the product of propagators associated with the graph $\Gamma$, and the numerator $n_\Gamma$ is all of the remaining kinematic factors including dot products of momenta and polarization vectors.

@@

Every numerator is required to respect all of the symmetries (or automorphisms) of its associated graph.
Graph symmetries are an avatar of Bose symmetry.
For a pure ansatz construction it is very reasonable to impose graph symmetries, but color-kinematics duality at tree-level is often achieved at the expense of manifest Bose symmetry \cite{Cheung:2016prv, Cheung:2021zvb, Brandhuber:2021bsf} \citepls{Alex numerators from CHY}.
In an ansatz construction graph symmetries are imposed for \emph{every} graph, not just the basis graphs.
%For example, the symmetries of the crossed box are spanned by
%\be
%\CrossedBoxGraphSym{1}{2}{3}{4} \sim \CrossedBoxGraphSym{2}{1}{3}{4} \text{ and } \CrossedBoxGraphSym{1}{2}{3}{4} \sim \CrossedBoxGraphSym{2}{1}{4}{3}
%\ee
For example, the symmetries of the crossed box
\be
\CrossedBoxGraphSym{1}{2}{3}{4}
\ee
are spanned by $\{1234\} \to \{2134\}$ and $\{1234\} \to \{1243\}$.
In general, internal edge labels must be tracked as well.
Furthermore, care must be taken with the signs in a symmetry relation.
To compensate for the antisymmetric color factors, $f^{abc}$, graph vertices are all totally antisymmetric and hence have signs built into their orientations.
Once the signs are included in the symmetry relations, the crossed box can be expressed in terms of the double box and penta-triangle, thus imposing non-trivial constraints on the ansatz.

@@

As one can see above, the Jacobi identity of \ref{cubicJac} fails only in terms with two factors of polarization dot products, $(\varepsilon\varepsilon)^2$. Whereas, if we restricted ourselves to just considering factors with a single polarization dot product, the jacobi identity of Yang-Mills would be satisfied \textit{off-shell} and in arbitrary spacetime dimensions:
\be
{}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4\big|^{(\varepsilon\varepsilon)^1}+\text{cyc}(234) =0
\ee
Thus for our approach at one-loop, we aim to build color-dual integrands directly from kinematic factors of this form. The organizational principle underlying this construction can be easily understood in terms of a polarization decomposition of vector amplitudes introduced by one of the authors \cite{Pavao:2022kog}, which we summarize briefly below. 
\subsection{Block Decomposition of Yang-Mills}
The $D$-dimensional Lagrangians of \ref{} can be understood as selecting specific on shell states to isolate the cubic sector of Yang-Mills at a given loop order. Start with reducible amplitude block decomp from Ref. \cite{Pavao:2022kog}:
\begin{equation}\label{eq:pureVecRABD}
A_{(\sigma)}^{\text{YM}} = \sum_{k=0}^{\lfloor |\sigma|/2\rfloor}\sum_{\rho \in S^{2|k}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\end{equation}
By a simple mass-dimensioin argument, we know the tree-level cubic sector must be at $\mathcal{O}((\varepsilon\varepsilon)^1)$:
\be
A_{(\text{tree})}^{\text{cubic}} = \sum_{\rho \in S^{2|1}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\ee
while at one-loop, the cubic sector corresponds to $\mathcal{O}((\varepsilon\varepsilon)^0)$ in polarization dot products:
\be
A_{(\text{1-loop})}^{\text{cubic}} = \sum_{\rho \in S^{2|0}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\ee
When decomposed in this way, in order for the full amplitude to be gauge invariant, the polarization stripped building blocks, must obey a set of Ward-Identities between different helicity sectors:
 \begin{equation}\label{eq:GIrelA}
\Delta_{(\sigma)}^{(\rho)}\Big|_{\epsilon_i\rightarrow k_i} =-
\sum_{j \in \rho^c} (k_i \varepsilon_j)\Delta_{(\sigma )}^{(\rho\cup (ij))} \,.
\end{equation}
With this in hand, we can construct a Lagrangian description of the of the cubic sector for Yang-Mills, and will demonstrate that the resulting amplitudes with manifestly color-dual Feynman rules are equivalent to both SDYM and NLSM through one-loop. 
\subsection{Semi-abelian Yang-Mills theory}
As we demonstrate above, if we include factors of $(\varepsilon\varepsilon)^{n\geq 1}$ at tree-level, then we need to keep the Yang-Mills four-point contact for color-kinematics duality to be restored on-shell. However, the contrapositive is also true -- if we get rid of the four-point Yang-Mills vertex, then we only need terms that contribute to the manifestly color-dual cubic sector of the theory.

@@

\subsection{One-loop color-dual integrands}
Our first application of this theory for color-dual construction at loop level is for self-dual Yang-Mills (SDYM). At tree-level, the off-shell cubic vertices can be written in terms of light cone coordinates \cite{}:
\be
X(p,k) = p_u k_w-p_w k_u
\ee
At loop-level, this construction needs to be analytically continued to general dimension in order to apply dimensional regularization at one-loop. This can be done using the cubic semi-YM vertices of the previous section:
\be
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= \mathcal{X}(k_1,k_2) = (\varepsilon_1\varepsilon_3)(\varepsilon_2 k_3) -  (\varepsilon_2\varepsilon_3)(\varepsilon_1 k_3)
\ee
Plugging in on-shell all-plus helicity states in light-cone gauge will yield precisely the 4D SDYM vertex, up to an unphysical phase. Of course, this has the advantage of permitting a $D$-dimensional construction of the one-loop integrand. Below we provide a four-point example:
\be
\simplebox = g_{\text{YM}}^4 \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\mathcal{X}(k_3,l_3)\mathcal{X}(k_4,l_4)\rangle
\ee
where $l_i = l+k_1+k_2+\cdots+ k_i$. In general, the $n$-gon vertex would be
\be
\mathcal{N}^{\text{SDYM}}_{n\text{-gon}}= \nGonVector = g_{\text{YM}}^n \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\cdots \mathcal{X}(k_n,l_n) \rangle
\ee
Due to the state-sum of internal loop factors, $\sum {\varepsilon_{(+l)}\varepsilon_{(-l)}}=D$, the integrand above depends explicitly on the spacetime dimension, $D$. By taking a derivative, we can recover the integrand needed for all-plus one-loop amplitudes with an internal scalar:
\be
\frac{\partial}{\partial D} \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}=\nGonScalar  = (2 g_{\text{YM}})^n (l_1 k_1)(l_2 k_2)\cdots (l_n k_n)
\ee
This integrand numerator with internal scalar loop is precisely what one would obtain from the Feynman rules of YZ-theory, absent the $XZ$ internal vector loop. For more background, we refer the reader to appendix \ref{}. 
\be
 \mathcal{N}^{\text{NLSM}}_{n\text{-gon}}=\left[\frac{\partial}{\partial D} \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}\right]^{\epsilon \rightarrow k}=\nGonScalar =  f_\pi^{-n} [12][23]\cdots [n1]
\ee
Where we have define the antisymmetric kinematic variable, $[LR] = l_L^2 - l_R^2$
\subsection{Two-loop obstruction}
At two-loop, introducing terms that conspire with four-point contacts is unavoidable (add diagram). Thus, to approach the problem of two-loop construction, we employ an ansatz based approach. 
\section{Two-loop bootstrap}

@@

Proceeding to the next-to-maximal cuts, we continue to avoid
pathological diagrams including the newly-appearing type discussed in
\cref{eq:other-path}. After discarding these types of cuts, we are
left with 8 \ace{(maybe a few more to account for dangling trees)}
unique cut topologies, shown in \cref{fig:ym-nmc}, to check.  Seven of
these next-to-maximal cuts are consistent with the kinematic Jacobi
relations and symmetries.  Interestingly, the ``kissing triangles''
cut,
\begin{equation*}
   \NMCutC \,,
\end{equation*}
cannot be satisfied by either the unrestricted or the BDN-restricted
ansatz with Jacobi relations and symmetries applied.
\begin{figure}
\centering
 { \NMCutH, \NMCutG,  \NMCutD,\NMCutB, 
 \\
 \NMCutE, \NMCutF, \NMCutA, \NMCutC,  }
  \caption{The 8 non-pathological next-to-maximal-cut diagrams.}
  \label{fig:ym-nmc}
\end{figure}

>>>>

Color-kinematics duality is a statement about the off-shell graphical simplicity that manifest redundancy in the on-shell observables of gauge and gravity theories. Ignoring factors of $i$ and the coupling constant, a generic $m$-point $L$-loop amplitude in a color-dual theory -- such as NLSM, YM, or bi-adjoint scalar theory -- can be written as
\be
\mathcal{A}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int \frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{c_\Gamma n_\Gamma}{d_\Gamma} .
\ee
Here the spacetime dimension is $D$.
Even though the underlying theory may have quartic (or higher multiplicity) vertices, it is possible to multiply and divide by propagators so that the sum runs over only purely cubic graphs $\Gamma$.
Overcounting is removed by the internal symmetry factor $S_\Gamma$, which tracks the number of graph symmetries with external legs held fixed.
The color factor $c_\Gamma$ is the string of structure constants $f^{abc}$ associated with the graph $\Gamma$,\footnote{This paper is focused on theories with particles that transform in the adjoint.  However, the double copy can be applied to a far more diverse set of theories with particles that transform in the fundamental or more diverse representations, see for example \cite{Johansson:2017srf, Johansson2014zca, Johansson:2015oia, Johansson:2019dnu, Carrasco:2023vjg, Carrasco:2020ywq} \jm{I cited DF$^2$, fundamental fermions, Aslan+JJ, Ingrid+JJ. Is there anything else anyone wants to mention?}.}
the denominator $d_\Gamma$ is the product of propagators associated with the graph $\Gamma$, and the numerator $n_\Gamma$ is all of the remaining kinematic factors including dot products of momenta and polarization vectors.

@@

Every numerator is required to respect all of the symmetries (or automorphisms) of its associated graph.
Graph symmetries are an avatar of Bose symmetry.
For a pure ansatz construction it is very reasonable to impose graph symmetries, but color-kinematics duality at tree-level is often achieved at the expense of manifest Bose symmetry \cite{Cheung:2016prv, Cheung:2021zvb, Brandhuber:2021bsf, Edison:2020ehu}.
In an ansatz construction graph symmetries are imposed for \emph{every} graph, not just the basis graphs.
%For example, the symmetries of the crossed box are spanned by
%\be
%\CrossedBoxGraphSym{1}{2}{3}{4} \sim \CrossedBoxGraphSym{2}{1}{3}{4} \text{ and } \CrossedBoxGraphSym{1}{2}{3}{4} \sim \CrossedBoxGraphSym{2}{1}{4}{3}
%\ee
For example, the symmetries of the crossed box
\be
\CrossedBoxGraphSym{1}{2}{3}{4}
\ee
are spanned by $\{1234\} \to \{2134\}$ and $\{1234\} \to \{1243\}$.
In general, internal edge labels must be tracked as well.
Furthermore, care must be taken with the signs in a symmetry relation.
To compensate for the antisymmetric color factors, $f^{abc}$, graph vertices are all totally antisymmetric and hence have signs built into their orientations.
Once the signs are included in the symmetry relations, the crossed box can be expressed in terms of the double box and penta-triangle, thus imposing non-trivial constraints on the ansatz.

@@

As one can see above, the Jacobi identity of \ref{cubicJac} fails only in terms with two factors of polarization dot products, $(\varepsilon\varepsilon)^2$. Whereas, if we restricted ourselves to just considering factors with a single polarization dot product, the Jacobi identity of Yang-Mills would be satisfied \textit{off-shell} and in arbitrary spacetime dimensions:
\be
{}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4\big|^{(\varepsilon\varepsilon)^1}+\text{cyc}(234) =0
\ee
Thus for our approach at one-loop, we aim to build color-dual integrands directly from kinematic factors of this form. The organizational principle underlying this construction can be easily understood in terms of a polarization decomposition of vector amplitudes introduced by one of the authors \cite{Pavao:2022kog}, which we summarize briefly below. 
\subsection{Block Decomposition of Yang-Mills}
The $D$-dimensional Lagrangians of \ref{} can be understood as selecting specific on shell states to isolate the cubic sector of Yang-Mills at a given loop order. Start with reducible amplitude block decomp from Ref. \cite{Pavao:2022kog}:
\begin{equation}\label{eq:pureVecRABD}
A_{(\sigma)}^{\text{YM}} = \sum_{k=0}^{\lfloor |\sigma|/2\rfloor}\sum_{\rho \in S^{2|k}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\end{equation}
By a simple mass-dimensioin argument, we know the tree-level cubic sector must be at $\mathcal{O}((\varepsilon\varepsilon)^1)$:
\be
A_{(\text{tree})}^{\text{cubic}} = \sum_{\rho \in S^{2|1}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\ee
while at one-loop, the cubic sector corresponds to $\mathcal{O}((\varepsilon\varepsilon)^0)$ in polarization dot products:
\be
A_{(\text{1-loop})}^{\text{cubic}} = \sum_{\rho \in S^{2|0}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\ee
When decomposed in this way, in order for the full amplitude to be gauge invariant, the polarization stripped building blocks, must obey a set of Ward-Identities between different helicity sectors:
 \begin{equation}\label{eq:GIrelA}
\Delta_{(\sigma)}^{(\rho)}\Big|_{\epsilon_i\rightarrow k_i} =-
\sum_{j \in \rho^c} (k_i \varepsilon_j)\Delta_{(\sigma )}^{(\rho\cup (ij))} \,.
\end{equation}
With this in hand, we can construct a Lagrangian description of the of the cubic sector for Yang-Mills, and will demonstrate that the resulting amplitudes with manifestly color-dual Feynman rules are equivalent to both SDYM and NLSM through one-loop. 
\subsection{Semi-abelian Yang-Mills theory}
As we demonstrate above, if we include factors of $(\varepsilon\varepsilon)^{n\geq 1}$ at tree-level, then we need to keep the Yang-Mills four-point contact for color-kinematics duality to be restored on-shell. However, the contrapositive is also true -- if we get rid of the four-point Yang-Mills vertex, then we only need terms that contribute to the manifestly color-dual cubic sector of the theory.

@@

\subsection{One-loop color-dual integrands}
Our first application of this theory for color-dual construction at loop level is for self-dual Yang-Mills (SDYM). At tree-level, the off-shell cubic vertices can be written in terms of light cone coordinates \cite{}:
\be
X(p,k) = p_u k_w-p_w k_u
\ee
At loop-level, this construction needs to be analytically continued to general dimension in order to apply dimensional regularization at one-loop. This can be done using the cubic semi-YM vertices of the previous section:
\be
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= \mathcal{X}(k_1,k_2) = (\varepsilon_1\varepsilon_3)(\varepsilon_2 k_3) -  (\varepsilon_2\varepsilon_3)(\varepsilon_1 k_3)
\ee
Plugging in on-shell all-plus helicity states in light-cone gauge will yield precisely the 4D SDYM vertex, up to an unphysical phase. Of course, this has the advantage of permitting a $D$-dimensional construction of the one-loop integrand. Below we provide a four-point example:
\be
\simplebox = g_{\text{YM}}^4 \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\mathcal{X}(k_3,l_3)\mathcal{X}(k_4,l_4)\rangle
\ee
where $l_i = l+k_1+k_2+\cdots+ k_i$. In general, the $n$-gon vertex would be
\be
\mathcal{N}^{\text{SDYM}}_{n\text{-gon}}= \nGonVector = g_{\text{YM}}^n \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\cdots \mathcal{X}(k_n,l_n) \rangle
\ee
Due to the state-sum of internal loop factors, $\sum {\varepsilon_{(+l)}\varepsilon_{(-l)}}=D$, the integrand above depends explicitly on the spacetime dimension, $D$. By taking a derivative\footnote{Another way to understand this derivative is that it selects the large $D$ behavior of the integrand.}, we can recover the integrand needed for all-plus one-loop amplitudes with an internal scalar:
\be
\frac{\partial}{\partial D} \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}=\nGonScalar  = (2 g_{\text{YM}})^n (l_1 k_1)(l_2 k_2)\cdots (l_n k_n)
\ee
This integrand numerator with internal scalar loop is precisely what one would obtain from the Feynman rules of YZ-theory, absent the $XZ$ internal vector loop. For more background, we refer the reader to appendix \ref{}. 
\be
 \mathcal{N}^{\text{NLSM}}_{n\text{-gon}}=\left[\frac{\partial}{\partial D} \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}\right]^{\epsilon \rightarrow k}=\nGonScalar =  f_\pi^{-n} [12][23]\cdots [n1]
\ee
Where we have define the antisymmetric kinematic variable, $[LR] = l_L^2 - l_R^2$
\subsection{Two-loop obstruction}
At two-loop, introducing terms that conspire with four-point contacts is unavoidable (add diagram). Thus, to approach the problem of two-loop construction, we employ an ansatz based approach. 
\section{Two-loop bootstrap}

@@

Proceeding to the next-to-maximal cuts, we continue to avoid
pathological diagrams including the newly-appearing type discussed in
\cref{eq:other-path}. After discarding these types of cuts, we are
left with 8 \ace{(maybe a few more to account for dangling trees)} \jm{How about we just say that there are 8 N$^1$ cuts ignoring dangling trees?}
unique cut topologies, shown in \cref{fig:ym-nmc}, to check.  Seven of
these next-to-maximal cuts are consistent with the kinematic Jacobi
relations and symmetries.  Interestingly, the ``kissing triangles''
cut,
\begin{equation*}
   \NMCutC \,,
\end{equation*}
cannot be satisfied by either the unrestricted or the BDN-restricted
ansatz with Jacobi relations and symmetries applied.
\begin{figure}
\centering
 { \NMCutH, \NMCutG,  \NMCutD,\NMCutB, 
 \\
 \NMCutE, \NMCutF, \NMCutA, \NMCutC,  }
  \caption{The 8 non-pathological next-to-maximal-cut diagrams.}
  \label{fig:ym-nmc}
\end{figure}
